\begin{prop}
\label{Prop 11}
Given a \(\Lambda\)-module \(A\) of finite length and a \(\Lambda\)-homomorphism \(f:A\mapsto A\). The following are equivalent
\begin{enumerate}[(a)]
    \item \(f\) is an isomorphism
    \item \(f\) is a monomorphism (1-1)
    \item \(f\) is an epimorphism (onto)
\end{enumerate}
\end{prop}
\begin{proof}
Clearly \((a)\implies (b)\) and \((a)\implies (c)\) by def. We have the exact sequence
\begin{align*}
    0\rightarrow f(A) \hookrightarrow A \rightarrow A/f(A) \rightarrow 0
\end{align*}
\noindent\underline{\((a)\implies (b)\)}: \(f\) 1-1 \(\implies A \simeq f(A) \implies l(A) = l(f(A))
\implies l(A/f(A)) = 0 \implies f(A) = A \implies f \text{  onto}.\)\\[0.3cm]
\noindent\underline{\((c) \implies (a)\)}: \(f\) onto \(\implies f(A) = A \implies l(A/f(A)) = 0 \implies l(A) = l(f(A))\\
c\rightarrow \Ker f \hookrightarrow A \rightarrow f(A) \rightarrow 0 \text{  exact} \implies l(\Ker f) = 0 \implies \Ker f = (0) \implies f \text{ 1-1 } \implies f \text{ isomorphism}.\)
\end{proof}
\begin{rem}
The proof of \ref{Prop 11} holds for all \(f:A\mapsto B\) with \(l(A) = l(B)\), so if \(l(A) = l(B)\) and \(f:A\mapsto B\), then 
\begin{align*}
    f \text{ isomorphism } \Leftrightarrow f \text{ 1-1 } \Leftrightarrow f \text{  onto}.
\end{align*}
\end{rem}

\noindent\textbf{Recall}: \(0\rightarrow A \rightarrow B \rightarrow C \rightarrow 0\) exact and \(l(B) < \infty\), then \(l(A)\) and \(l(C)\) are finite too.
\begin{prop}
\label{Prop 12}
If \(\xymatrix{0\ar[r] & A\ar[r]^f & B\ar[r]^g & C\ar[r] & 0}\) is exact and \(A\) and \(C\) have finite length, then also \(B\) has finite length and \(l(B) = l(A) + l(C)\).
\end{prop}
\begin{proof}
Let 
\begin{align*}
    \mathcal{F}': A = A_0 \supseteq A_1 \supseteq\dots\supseteq A_{n-1}\supseteq A_n = (0)
\end{align*}
and
\begin{align*}
    \mathcal{F}'': C = C_0 \supseteq C_1 \supseteq\dots\supseteq C_{m-1}\supseteq C_m = (0)
\end{align*}
be two comp. series of \(A\) and \(C\), respectivley. Consider the following chain of submods of \(B\):\\[0.2cm]
\begin{align*}
    \mathcal{F}: B = g^{-1}(C) \supseteq g^{-1}(C_1) \supseteq g^{-1}(C_2) \supseteq\dots\supseteq g^{-1}(C_m) = \Ker g
\end{align*}
and
\begin{align}
    \Ker g = \Im f = f(A) \subseteq f(A_1) \subseteq f(A_2) \subseteq\dots\subseteq f(A_{n-1}) \subseteq f(A_n) = (0)
\end{align}
We want to show that \(\mathcal{F}\) is a comp. series of \(B\). Let \(g_i = g|_{g^{-1}C_i}: g^{-1}(C_i)\mapsto C_i \big(b_i\mapsto g(b_i)\big)\). Then \(g_i\) is clearly surjective (since \(g\) is). The composition
\begin{align*}
\xymatrix{\Psi_i:\Pi_i g_i: g^{-1}(C_i)\ar[r] & C_i\ar[r]^{\Pi_i} & C_i/C_{i+1}  \\
 & & c_i \mapsto c_i + C_{i+1}}
\end{align*} 
 onto (comp. of two onto maps) and we have \(b_i\in \Ker \Psi_i \Leftrightarrow \Pi_i(g_i(b_i)) = 0\) in \(C_i/C_{i+1} \Leftrightarrow g(b_i) + C_{i+1} = 0\) so \(g(b_i))\in C_{i+1} \Leftrightarrow b_i\in g^{-1}(C_{i+1})\implies \Ker\Psi_i = g^{-1}/C_{i+1})\implies g^{-1}(C_i)/g^{-1}(C_{i+1}) \simeq \Im \Psi_i\) since \(g^{-1}(C_{i+1}) = \Ker \Psi_i\) and \(C_i/C_{i+1}\) is simple by definition of \(\mathcal{F}''\)
 \end{proof}
 \\[0.5cm]
 \hline
 .\\[0.5cm]
Let \(f_i = f|_{A_i}:A_i\rightarrow f(A_i\) (i.e. \(a_i\mapsto f(a_i)\)) which clearly is onto. The composition
\begin{align*}
    \xymatrix{\theta_i = p_if_i:A_i\ar[r]^{f_i} & f(A_i)\ar[r]^{p_i} & f(A_i)/f(A_{i+1})}
\end{align*}
where \(p_i(b_i) = b_i + f(A_{i+1})\), is onto (composition of two onto maps) and we have
\begin{gather*}
    a_i\in\Ker \theta_i \Leftrightarrow p_if_i(a_i) = 0 \text{ in } f(A_i)/f(A_{i+1})\\
    \Leftrightarrow f(a_i) + f(A_{i+1}) = 0\\
    \Leftrightarrow f(a_i)\in f(A_{i+1})\\
    \exists a_{i+1}\in A_{i+1}\text{ such that } f(a_{i+1} = f(a_i) \text{   (because } f 1-1)\\
    \Leftrightarrow a_i = a_{i+1}\\
    \Leftrightarrow a_i\in A_{i+1}\\
    \implies \Ker\theta_i = A_{i+1}\\
    \implies A_i/\Ker\theta_i \simeq \Im\theta_i = f(A_i)/f(A_{i+1})
\end{gather*}
\(A_i/\Ker\theta_i = A_i/A_{i+1}\) is simple by definition of \(\mathcal{F}'\). Hence \(\mathcal{F}\) is a composition series of \(B\) and
\begin{align*}
    l(B) = l(A) + l(C)
\end{align*}
\hline
.\\[0.5cm]
\begin{defin}
    A collection \(\mathcal{C}\) of modules (a full subcategory) is \underline{closed under extensions} if for each exact sequence 
    \begin{align*}
        \xymatrix{0\ar[r] & A\ar[r] & B\ar[r] & C\ar[r] & 0}
    \end{align*}
    with \(A, C\in\mathcal{C}\), then \(B\) is also in \(\mathcal{C}\)
\end{defin}
\index{Closed under extensions}
Let \(fl(\Lambda\) be the collection of \(\Lambda\)-modules of finite length.
\begin{prop}
\label{Proposition 13}
\begin{enumerate}[(a)]
    \item \(fl(\Lambda)\) is closed under extensions and contains the simples. Furthermore, \(fl(\Lambda)\) is closed under submodules and factor modules.
    \item Let \(\mathcal{C}\) be a collection of \(\Lambda\)-modules that is closed under extensions and contains the simple \(\Lambda\)-modules. Then \(fl(\Lambda)\subseteq\mathcal{C}\)
\end{enumerate}
\end{prop}
\begin{proof}(a): [FIX REFERENCES]\\
(b) Let \(B\in fl(\Lambda)\) with \(l(B) = n\). Induction on \(n\).\\
\noindent \(n = 1\): Then \(B\) is simple and \(B\in\mathcal{C}\).\\
\noindent\(n>1\): Choose \(0\not\subseteq A\not\subseteq B\) submodule. This is possible since \(B\) is not simple. Then
\begin{align*}
    \xymatrix{0\ar[r] & A\ar[r] & B\ar[r] & C\ar[r] & 0}
\end{align*}
is exact with \(l(A),l(B/A)<l(B) = n\). Induction \(\implies A, B/A\in\mathcal{C}\). \(\mathcal{C}\) is closed under extensions so \(B\in\mathcal{C}\) and hence \(fl(\Lambda) \subseteq \mathcal{C}\).
\end{proof}
\noindent\underline{Recall:} A module \(M\) is Noetherian (Artinian) if for every ascending (descending) chain of submodules of M:
\begin{gather*}
    M_1\subseteq M_2\subseteq\dots\subseteq M_n \subseteq M_{n+1} \subseteq\dots\subseteq M \\
    (M \supseteq M_1\supseteq M_2\supseteq\dots\supseteq M_n \supseteq M_{n+1} \supseteq\dots)
\end{gather*}
\(\exists n\) such that \(M_n = M_{n+1} = \dots\).

\noindent \(M\) is Noetherian (Artinian) if, and only if, every non-empty set of sub-modules of \(M\) has a maximal (minimal) element.

\begin{prop}
A \(\Lambda\)-module. \(l(A) < \infty \Leftrightarrow\) A is  Artinian and Noetherian.
\end{prop}
\begin{proof}\underline{\(\implies\):} Want to show: \(fl(\Lambda) \subseteq \text{art}(\Lambda) = \) collection of Artinian \(\Lambda\)-modules. Will use Proposition \ref{Proposition 13} (b). Let 
\begin{align*}
    \xymatrix{0\ar[r] & A\ar[r] & B\ar[r] & C\ar[r] & 0}
\end{align*}
be exact. We claim that \(A, C\in\text{art}(\Lambda) \implies B\in\text{art}(\Lambda)\). Consider a descending chain \( B = B_0 \subseteq B_1 \subseteq \dots\). As before we get induced chains of submodules in \(A\) and \(C\). 
\begin{gather*}
    A = A_0 \supseteq A_1 \supseteq\dots\supseteq A_i\supseteq\dots\quad\text{with } A_i = f^{-1}(B_i)\\
    C = C_0 \supseteq C_1 \supseteq\dots\supseteq C_i\supseteq\dots\quad\text{with } C_i = g(B_i)
\end{gather*}
and induced exact squence
\begin{align*}
    \xymatrix{0\ar[r] & A_i/A_{i+1}\ar[r]^{\bar{f}} & B_i/B_{i+1}\ar[r]^{\bar{g}} & C_i/C-{i+1}\ar[r] & 0}
\end{align*}
since \(A, C\in\text{art}(\Lambda)\) there exists \(N\) such that
\begin{gather*}
A_i = A_{i+1} \text{ and } C_i = C_{i+1}\quad i\geq N\\
\implies A_i/A_{i+1} = (0) \text{ and } C_i/C_{i+1} = (0) \quad i \geq N \\
\implies B_i/B_{i+1} = 0 \implies B_i = B_{i+1} \quad i\geq N\\
\implies B\in\text{art}(\Lambda) \\
\implies \text{art}(\Lambda) \text{ is closed under extensions.}
\end{gather*}
Clear that \(\text{art}(\Lambda)\) contains the simple \(\Lambda\)-modules: Proposition \ref{Proposition 13}(b)\(\implies  \text{fl}(\Lambda)\subseteq\text{art}(\Lambda)\)

\noindent\underline{Exercise:} Similarly, \(\text{fl}(\Lambda)\subseteq\text{noeth}(\Lambda)\), collection of noehterian \(\Lambda\)-modules. This implies that \(\text{fl}(\Lambda)\subseteq\text{art}(\Lambda)\cap\text{noeth}(\Lambda)\).\\[0.5cm]

\noindent\underline{\Large\(\Leftarrow\)}: Assume that \(B\neq (0)\) is Artinian and Noetherian. Since \(B\) is Artinian \(B\) has a simple submodule \(S\subseteq B\). (\(\mathcal{F} = \{U\subseteq B\,\, |\,\, U\neq (0)\}\) has a minimal element). Consider \(\mathcal{F}' = \{U\subseteq B\,\, |\,\, l(U) < \infty \}\). Then \(\mathcal{F}' \neq \emptyset\) since \(S\in\mathcal{F}'\). Since \(B\) is Noetherian \(\mathcal{F}'\) has a maximal element \(A\subseteq B\) and \(A \in \text{fl}(\Lambda)\). Assume that \(A\subsetneq B\), i.e. \(B/A\neq (0)\). \(B\) Artinian \(\implies B/A\) Artinian \(\implies \exists T\subseteq B/A\) simple submodule. Consider the natural projection \(p:B\rightarrow B/A\). Then 
\begin{align*}
    p|_{p^{-1}} :p^{-1}(T)\rightarrow T \quad(\subseteq B/A)
\end{align*}
is onto and \(\Ker p|_{p^{-1}} = A\). Hence \((p^{-1}(T))/A\simeq T\) and we have an exact sequence
\begin{align*}
    \xymatrix{0\ar[r] & A\ar[r] & p^{-1}(T)\ar[r] & T\ar[r] & 0}
\end{align*}
Now \(l(p^{-1}(T)) = l(A) + 1\), a contradiction and we conclude that \(A = B\) and \(B\in\text{fl}(\Lambda)\).
\end{proof}
\noindent\underline{Note: } 
\begin{enumerate}[(1)]
    \item If \(\Lambda\) is a ring with 1 then \(\Lambda\) is artinian if, and only if, \(l(_{\Lambda}\Lambda)<\infty\)\\
    \begin{proof} \underline{\(\Rightarrow\)}: \(\Lambda\) left Artinian \(\implies \Lambda\) left Noetherian \(\implies _\Lambda\Lambda\in\text{art}(\Lambda)\cap\text{noeth}(\Lambda)\implies l(_\Lambda\Lambda)<\infty\)\\
    \underline{\(\Leftarrow: \)} \(l(_\Lambda\Lambda) < \infty \implies _\Lambda\Lambda \in \text{art}(\Lambda)\cap\text{noeth}(\Lambda)\implies\Lambda\) left Artinian.
    \end{proof}
    \item \(\Lambda\) is left Artinian. \textit{Challenge:} \(\text{fl}(\Lambda) = \text{mrd(?)}(\Lambda)\) - finitely generated \(\Lambda\)-modules. 
    \item \(\Lambda = \mathbb{Z}, M = \mathbb{Z}/(n)\) \\
    \(n = p_1^{m_1}p_2^{m_2}\dots p_t^{m_t}\quad,\quad p_i\) different primes and \(m_i\geq 1\)\\
    \(m_{\mathbb{Z}/p}(M) = \begin{cases}m_i & \text{if } p = p_i \\ 0 & \text{otherwise}\end{cases}\)
\end{enumerate}
\begin{prop}
    \(\Lambda\) a ring, \(B\) semisimple \(\Lambda\)-module. TFAE:
    \begin{enumerate}[(a)]
        \item \(l(B) <\infty \)
        \item \(B\) is Artinian. 
        \item \(B\) is Noetherian
    \end{enumerate}
\end{prop}
\begin{proof}
    Excercise
\end{proof}
